---
title: "R Basics In Class Practice - Solutions"
output: html_notebook
---
Covered topics:
Basic Data Types
Basic Arithmetic
Vector
Vector Operations
Comparison operators
Vector Selection and Indexing


Solve the following problems using R - What is two to the power of five? 
```{r}
2^5
```

Create a vector called stock.prices with the following data points: 23,27,23,21,34 
```{r}
stock.prices <- c(23,27,23,21,34)
stock.prices
```

Assign names to the price data points relating to the day of the week, starting with Mon, Tue, Wed, etc...

```{r}
names(stock.prices) <- c('Mon','Tues','Wed','Thu','Fri')
stock.prices
```


What was the average (mean) stock price for the week? (You may need to reference a built-in function) 
```{r}
mean(stock.prices)
```


Create a vector called over.23 consisting of logicals that correspond to the days where the stock price was more than $23 

```{r}
over.23 <- stock.prices>23
```


Use the over.23 vector to filter out the stock.prices vector and only return the day and prices where the price was over $23 
```{r}
stock.prices[over.23]
```

Use a built-in function to find the day the price was the highest
```{r}
max(stock.prices)
```

```{r}
max.price <- stock.prices == max(stock.prices)
max.price
```
```{r}
stock.prices[max.price]
```
