---
title: "Week 5 Assignment"
author: "Oluwafemi Fabiyi"
date: "11/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading relevant libraries

```{r cars 1}
# install.packages("readr")
# install.packages("tidyverse") # include magrittr
# install.packages("stringr")
# install.packages("lattice")
# install.packages("ggplot2")

library(readr)
library(tidyverse)
library(stringr)
library(lattice)
library(ggplot2)
```

###### 1)	Create a function to calculate a*b+log(a); use it to get calc(3,4) and calc(5,6)

```{r cars 1.1}

calcLog <-function(a, b){
  return (a*b+log(a))
}
```
###### 2)	Create a function to check if the number is divisible by three or not

```{r cars 1.2}

divBy3 <- function(val){
  result <- FALSE
  if(val %% 3 == 0){
    result == 3
  }
  return (result)
}
```
###### 3) Create a list with the following 5 dataframes: 

```{r cars 1.3}
Data1<-data.frame("State"=c("CA","NY","IL"),"Value"=c(900,1000,1200))
Data2<-data.frame("State"=c("MN","WI","IN"),"Value"=c(300,800,700))
Data3<-data.frame("State"=c("AZ","AR","TX"),"Value"=c(200,500,1300)) 
Data4<-data.frame("State"=c("MD","NJ","WA"),"Value"=c(400,1100,600))
Data5<-data.frame("State"=c("MA","NC","AL"),"Value"=c(100,1600,300))

myList <- list(Data1, Data2, Data3, Data4, Data5)
myList
```
###### 4) Create a function using a for loop to stack the 5 Dataframes (add rows) in the list from questions 2:

```{r cars 1.4}
sizeOfOriginalList <- length(myList)
sizeOfOriginalList
for(i in 1:sizeOfOriginalList){
newIndex <- i + sizeOfOriginalList
myList[newIndex] <- myList[i]
}
myList
```
###### 5.1) GL Data Exploration (10 points)
Display the data structure and overview of the rows and columns of the dataset
Use the summary method to display descriptive statistics for every variable

```{r cars 5.1}
dataPath <- "~/Desktop/MSc Analytics/Autumn 2020/MSCA 37010 2 - Programming for Analytics/Week 5/Assignment/"
dat <- read.csv(paste(dataPath,'GeneralLiabilityExposure.csv',sep = '/'), header=TRUE)
head(structure(dat))
summary(dat)
```
###### 5.2) Create a Scatterplot for the Exposure column (5 points)

```{r cars 5.2}
plot(dat$Exposure)
```

###### 5.3) Create a Histogram of the Premium column using your favorite color (5 points)

```{r cars 5.3}
hist(dat$Premium, 
     main = 'Premium Values',xlab = 'Premium', 
     col='blue')
```

###### 6.1 Use the ClaimSeverity column in the AutoCollision dataset to create a horizontal plot (5 points)
```{r cars 6.1}
dataPath <- "~/Desktop/MSc Analytics/Autumn 2020/MSCA 37010 2 - Programming for Analytics/Week 5/Assignment/"
mydat <- read.csv(paste(dataPath,'AutoCollision.csv',sep = '/'), header=TRUE)
head(structure(mydat))
summary(mydat)


barplot(mydat$ClaimSeverity, 
        main = 'Claim Severity Graph',
        xlab = 'calim severity levels', 
        col= 'green',horiz = TRUE)
```

###### 6.2) Use the ClaimSeverity column in the AutoCollision dataset to create a vertical plot (5 points)

```{r cars 6.2}
barplot(mydat$ClaimSeverity, 
        main = 'Claim Severity Graph',
        xlab = 'calim severity levels', 
        col= 'green',horiz = FALSE)
```

###### 6.3) Use the ClaimCount column in the AutoCollision dataset to create a Single box plot (5 points)

```{r cars 6.3}
#Single box plot
boxplot(mydat$ClaimCount, main='Single Boxplot')
```

###### 6.4) Use the 15-17 columns in the GL dataset to create a Multiple box plot (10 points)

```{r cars 6.4}
# Multiple box plots
boxplot(dat[,15:17], main='Multiple Boxplots')

```

###### 6.5) Put all the plots above in a single plot (5 points)

```{r cars 6.5}
par(mfrow=c(3,3), mar=c(2,5,2,1), las=1, bty="n")
barplot(mydat$ClaimSeverity, 
        main = 'Claim Severity Graph',
        xlab = 'calim severity levels', 
        col= 'green',horiz = TRUE)
barplot(mydat$ClaimSeverity, 
        main = 'Claim Severity Graph',
        xlab = 'calim severity levels', 
        col= 'green',horiz = FALSE)
boxplot(mydat$ClaimCount, main='Single Boxplot')
boxplot(dat[,15:17], main='Multiple Boxplots')
```


###### 7) Use the AutoCollision dataset to create a plot for each of ClaimSeverity and ClaimCount 
columns arranged in a Scatterplot Matrix (Hint: use splom())
```{r cars 7.1}
splom(mydat[c(3,4)], main="severity / count")
```

###### 8) Use R built-in dataset mtcars to create a styling scatter plots with disp as x, hp as y, cyl as factor
```{r cars 8.0}
library(ggplot2)
attach(mtcars)
palette("default")
cyl <- factor(cyl,levels=c(4,6,8),  
              labels=c("4cyl","6cyl","8cyl"))
ggplot(data = mtcars, mapping = aes(x = disp, y = hp, color = am)) + geom_point()

```


###### 9.1)	What is the function to remove object x?
```{r cars 9.1}
x <- c(1:3)
x
rm(x)

```

###### 9.2)	What is the garbage collection function used to releases memory, freeing it up to used once again?
```{r cars 9.2}
 gc()
```

###### 9.3)	What function is used to check how much memory is being used?
```{r cars 9.3}
 memory.profile()
```



