---
title: "Assignment 2 - MSCA 37010 2 Programing for Analytics_Oluwafemi Fabiyi"
author: "Oluwafemi Fabiyi"
date: "10/16/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1

Create 2 vectors A and B, where A is (1,2,3,4) and B is (5,6,7,8). You'll need to choose the cbind() or rbind() function correctly to create a 2 by 4 matrix from the vectors. (5 points)

```{r 1}
A <- c(1,2,3,4)
B <- c(5,6,7,8)

rbind(A,B)
```

### 2

2.1) Create a 4 by 4 matrix consisting of the numbers 1-16. Create this matrix using 1:16 in the matrix() function call. Assign this matrix to the variable Mat (10 points)

```{r 2}
Mat <- matrix(1:16, byrow = TRUE, nrow = 4, ncol = 4)

```

 2.2) Confirm that Mat is a matrix
 
```{r 2.2}
is.matrix(Mat)
```

2.3) The top row should be the numbers 1-4

```{r 2.3}
Mat
```

2.4) Using indexing notation, grab a sub-section of Mat from the previous exercise that looks like this:
 [11,12]
 [15,16]
```{r 2.4}
Mat[3:4, 3:4]
```

2.5) What is the sum of all the elements in Mat?
```{r 2.5}
sum(Mat)
```

###  3
Use runif() function to create a 4 by 5 matrix consisting of 20 random numbers between 0 and 100. (10 points)
Note: The runif() function generates random deviates about the uniform distribution on the interval from min to max
Hint: use ?runif

```{r 3}
matrix(runif(20, 0, 100), nrow = 4, ncol=5)
```

###  4
(10 points)
```{r 4}
y<- 1:3; y
A= matrix(1:9, nrow=3)
B= matrix(15:23,nrow=3)

```
             
4.1)	Calculate math expression multiplication
```{r 4.1}
A*B
```
4.2)	Calculate matrix multiplication
A%*%B
```{r 4.2}
A%*%B
```

4.3)  Write code to calculate A divided by y%*%y
```{r 4.3}
A/c(y%*%y)
```

###  5
(10 points)
      
 5.1)	Grab the elements of row 1 and column 3 
 
```{r 5.1}
x <- matrix(c(0,6,1,3,5,9), nrow=2, ncol=3, 
             dimnames=list(rows=c("r.1", "r.2"),cols=c("c.1", "c.2", "c.3")))       
x[1,3]

```
 
 5.2)  Grab the elements of third column   
```{r 5.2}
x[,3]

```
 5.3)	Grab Rows 1 and 2, all Columns
```{r 5.3}
x[,3]

```
### 6
Use apply() function and x <- matrix(1:9, nrow=3, ncol=3) to calculate: (10 points)                                    
6.1)	Row mins
```{r 6.1}
x <- matrix(1:9, nrow=3, ncol=3)
x
apply(x,1,min)

```
6.2)	Column sums

```{r 6.2}
apply(x,2,sum)

```

### 7 

7.1) Use the factor() function and degrees to create ordinal categorical variables. (10 points)

```{r 7.1}
degrees <- c('BSc','MSc','PhD','BSc','BEng','MEng','PhD','BSc','BSc','MSc','BEng')
fact.degrees <- factor(degrees); 
fact.degrees

```

 7.2) Summarize the numbers by different degrees

```{r 7.2}
summary(fact.degrees)

```

### 8

Use apply() function and y <- array(1:12,dim =c(2,2,3), dimnames=list(c("Row1","Row2"), c("Column1","Column2"), c("Matrix1","Matrix2","Matrix3")))  to calculate:  (10 points)                                                         
8.1)	 Column sums
```{r 8.1}
y <- array(1:12,dim =c(2,2,3), dimnames=list(c("Row1","Row2"), c("Column1","Column2"), c("Matrix1","Matrix2","Matrix3")))
y
apply(y, 2, sum)
```
8.2)	 Row means
```{r 8.2}
apply(y, 1, mean)
```

### 9

Use drink <- c("redwine", "whitewine", "water")  (10 points)     

9.1) What is the number of strings of drink?
```{r 9.1}
drink <- c("redwine", "whitewine", "water") 
length(drink)

```

9.2) What is the number of characters in each string?
```{r 9.2}
nchar(drink)

```

9.3)	What are the characters 2-4 of each drink?
```{r 9.3}
substring(drink, 2, 4)

```

9.4)	How to print the first three characters of drink?
```{r 9.4}
substr(drink, 1, 3)
```

9.5)	How to print drink in all upper case?
```{r 9.5}
toupper(drink)

```

### 10

Use the following codes to import KaggleHealthcareNoshows.csv (15 points)

```{r 10}
#install.packages("readr")
library(readr)
getwd()
setwd('/Users/oluwafemifabiyi')
khc<- read_csv('KaggleHealthcareNoshows.csv')
#View(khc)

```

10.1) View the structure of data (2.5 points)

```{r 10.1}
str(khc)

```

10.2) List the column names (2.5 points)
```{r 10.2}
colnames(khc)

```
 

10.3) First 5 Row names (2.5 points)

```{(r 10.3}
head(rownames(khc),5)

```

10.4) Dimension of the dataframe (2.5 points)
```{r 10.4}
dim(khc)

```

10.5) Create a data set with "PatientId", "AppointmentID","Gender" columns (2.5 points)
```{r 10.5}
khc[ , c('PatientId','AppointmentID','Gender')]

```

10.6) Use two ways to pull No.show column (2.5 points)
```{r 10.6}
khc[,14]
khc$'No-show'

```

10.7) Use factor() function to summarize the count of No.show column by Yes and No (5 bonus points)

```{r 10.7}
fact.No.show <- factor(khc$'No-show',ordered=TRUE,levels=c('No','Yes'))
summary(fact.No.show)

```

10.8) Calculate Average Age by No.show = Yes and No. (5 bonus points)


```{r 10.8}
#install.packages("magrittr") # package installations are only needed the first time you use it
#install.packages("dplyr")    # alternative installation of the %>%
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)    # alternatively, this also loads %>%

aggregate(khc$Age,by=list(khc$'No-show'),FUN = mean)

```
