---
title: "Week 4 Assignment"
author: "Oluwafemi Fabiyi"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1

```{r cars 1}
library(readr)
library(dplyr)
library(MASS)

head(painters, 3)
```
###### 1.1)	To select the observations of painters with Drawing greater than or equals 18
```{r cars 1.1}

head(filter(painters,Drawing >= 18))

```
###### 1.2)	To select those observations of School is B and D
```{r cars 1.2}

head(filter(painters, School == 'B' | School == 'D'))

```
###### 1.3)	To select those observations with the Composition (the first column) and Expression variables (the fifth column)
```{r cars 1.3}

head(dplyr::select(painters,Composition, Expression))

```
###### 1.4)  To create a data set that contains those observations of School is B and D with the Composition (the first column) and Expression variables (the fifth column)
```{r cars 1.4}

painters %>% dplyr::select(School, Composition, Expression) %>% filter(School == 'B' | School == 'D')

```
### 2 
Use CommercialProperty.csv to do the following: (20 points)

```{r cars 2}
dataPath <- "~/Desktop/MSc Analytics/Autumn 2020/MSCA 37010 2 - Programming for Analytics/Week 4/Assignment/"
dat <- read.csv(paste(dataPath,'CommercialProperty(2).csv',sep = '/'), header=TRUE)
head(dat, 3)

```
###### 2.1) Subset/Create a EffectiveYear variable based on the EffectiveDate column
```{r cars 2.1}
EffectiveYear <- format.Date(as.Date(dat$EffectiveDate,format="%d-%b-%y"), "%Y")
head(EffectiveYear)
```

###### 2.2) Subset/Create a EffectiveMonth variable based on the EffectiveDate column
```{r cars 2.2}
EffectiveMonth <- format.Date(as.Date(dat$EffectiveDate,format="%d-%b-%y"), "%B")
head(EffectiveMonth)
```

###### 2.3) Subset/Create a dataset with EarnedPremium>=3000 and EffectiveYear>=17
```{r cars 2.3}
dataset2.3 <- dat %>% filter(EarnedPremium>=3000)
head(dataset2.3)

```

###### 2.4) Subset/Create a dataset with IncurredLoss>100 
```{r cars 2.4}
dataset2.4 <- dat %>% filter(IncurredLoss>100)
head(dataset2.4)
```

###### 2.5) Subset/Create a sub dataset with State is CA and Segment is habitational
```{r cars 2.5}
dataset2.5 <- dat %>% filter(State == 'CA')
head(dataset2.5)
```


###### 2.6) Subset/Create a dataset when IncurredLoss>100 and State is CA and Segment is Habitational
```{r cars 2.6}
dataset2.6 <- dat %>% filter(IncurredLoss>100) %>% filter(State == 'CA')
head(dataset2.6)

```

###### 2.7) Calculate average earned premium:
```{r cars 2.7}
mean(dat$EarnedPremium)
```

###### 2.8) Create a variable named as "LargePremium_Indicator" when the EarnedPremium is greater than average EarnedPremium from the above
```{r cars 2.8}
LargePremium_Indicator <- dat %>% filter(EarnedPremium > mean(dat$EarnedPremium))
head(LargePremium_Indicator)
```

###### 2.9) Use for loop to create an indicator (aka: binary variable/new column) for each of CA, IL, FL, NY based on the State column # (10 Bonus Points)
```{r cars 2.9}
bin_value <- c()
for(val in 1:length(dat$State)){
  if(dat$State[val] == 'CA'){
    bin_value[val] <- 1
  }
  else if (dat$State[val] == 'IL'){
      bin_value[val] <- 1
  }
  else if (dat$State[val] == 'FL'){
      bin_value[val] <- 1
  }
  
  else if (dat$State[val] == 'NY'){
      bin_value[val] <- 1
  }
  else{
      bin_value[val] <- 0
  }
}
dat$new_col <- bin_value
head(dat)

```

### 3
```{r cars 3}

val <- 'B-'
if(val == 'A'){
  print('93%–100%')
}else if(val == 'A-'){
    print('90%–92%')
}else if(val == 'B+'){
    print('87%–89%')
}else if(val == 'B'){
    print('83%–86%')
}else if(val == 'B-'){
    print('80%–82%')
}else if(val == 'C+'){
    print('77%–79%')
}else if(val == 'C'){
    print('73%–76%')
}else if(val == 'C-'){
    print('70%–72%')
}else if(val == 'D'){
    print('60%–69%')
}else{
    print('0% - 59%')
}

```

### 4
Import the CommercialProperty dataset and Attach the CommercialProperty dataset (35 points)
```{r cars 4}
CP <- "~/Desktop/MSc Analytics/Autumn 2020/MSCA 37010 2 - Programming for Analytics/Week 4/Assignment/"
CP <- read.csv(paste(dataPath,'CommercialProperty.csv',sep = '/'), header=TRUE)
names(CP)
```

###### 4.1) Use a function to signal an error if the CommercialProperty dataset contains NA
```{r cars 4.1}
if(any(is.na(CP)) == TRUE){
  print('Error')
}

```

###### 4.2) Use a function to remove missing data from the CommercialProperty dataset 
```{r cars 4.2}
cleanCP <- na.omit(CP)
head(cleanCP)
```

###### 4.3) Use a function to return a logical vector indicating which rows have no missing data
```{r cars 4.3}
head(complete.cases(CP))
is.vector(complete.cases(CP))
typeof(complete.cases(CP))
```

###### 4.4) Sort the columns by PolicyNumber and EarnedPremium in ascending order
```{r cars 4.4}
head(CP %>% arrange(PolicyNumber,EarnedPremium))

```

###### 4.5) Use a ifelse() function to create an indicator for each of CA, IL, FL, NY
```{r cars 4.5}
  if(dat$State[1] == 'CA'){
    print('California')
  }else if (dat$State[1] == 'IL'){
      print('Illinois')
  }else if (dat$State[1] == 'FL'){
      print('FLorida')
  }else if (dat$State[1] == 'NY'){
      print('New York')
  }else{
      print('Other')
  }

```

###### 4.6) Create a variable LossRatio (= IncurredLoss/EarnedPremium) 
```{r cars 4.6}

```

###### 4.7) Create an indicator (= 1 if Loss Ratio >=60% & IncurredLoss >=25,000, else 0)
```{r cars 4.7}
# - --- dat = dat %>% mutate(candy.flag = factor(ifelse(var == "Candy", "Candy", "Non-Candy")))

```

# 5
Use For Loop to calculate squared values for 1 to 25.(5 points)
```{r cars 5}
for(val in 1:25){
  print(val * val)
}

```

# 6
Use the Workouts vector to create a for loop to list all the weekdays (5 points)
Workouts <- c("Biking","Hiking","Yoga","Walking","Jogging")

######  Hint: Some possible output looking like below:

######  [1] "Biking"  "on"  "Monday"
######  [1] "Hiking"  "on"  "Tuesday"
######  [1] "Yoga"    "on"  "Wednesday"
######  [1] "Walking" "on"  "Thursday"
######  [1] "Jogging" "on"  "Friday"
```{r cars 6}
Workouts <- c("Biking","Hiking","Yoga","Walking","Jogging")
WeekDays <- c("Monday","Tuesday","Wednesday", "Thursday", "Friday")

for(val in 1: length(Workouts)){
  cat(Workouts[val],' on ',WeekDays[val],'\n')
}

```

# 7
Subset/Create a for loop to convert fahrenheit to Celeius in the vector: (5 point)
###### Hint: use length() function to count the temperature position in fahrenheit vector
```{r cars 7}
fahrenheit <- c(60, 65, 70, 75, 80, 85, 90, 95, 100) 
for(val in 1: length(fahrenheit)){
  c = (fahrenheit[val]-32) * 5/9
  cat(fahrenheit[val],'F', ' is equal to ', c, '\n')
}

```
# 8
Use a for loop to count the number of even numbers in a vector. (5 points)

```{r cars 8}
x <- c(2,5,3,9,8,11,6)
i <- 0
for(val in 1: length(x)){
  if(x[val] %% 2 == 0){
    i = i + 1
  }
}
print(i)
```
# 9
Please create a for loop to merge the 5 Dataframes (add columns): (10 bonus points)
```{r cars 9}
Data1<-data.frame("State"=c("CA","NY","IL"),"Value1"=c(900,1000,1200))
Data2<-data.frame("State"=c("CA","NY","IL"),"Value2"=c(300,800,700)) 
Data3<-data.frame("State"=c("CA","NY","IL"),"Value3"=c(200,500,1300))
Data4<-data.frame("State"=c("CA","NY","IL"),"Value4"=c(400,1100,600))
Data5<-data.frame("State"=c("CA","NY","IL"),"Value5"=c(100,1600,300)) 

my_list <- list(Data1, Data2, Data3, Data4, Data5)

for(val in 1:length(my_list)){
  i = val + 1
  Data1[,i] <- my_list[[val]][2]
}

Data1

```
