---
title: "Assignment Week 3"
author: "Oluwafemi Fabiyi"
date: "10/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1

Please submit your R script with output for the following questions.(10 points)
1) Given the following two vectors:
x <- 1:15; y <- 5:25

1.1) What elements of x are in y?
```{r 1.1}
x <- 1:15
y <- 5:25

x
y
```

1.2) What is the output of x%%3?
```{r 1.2}
x %% 3

```

1.3) Use a function to see is there any x greater than 5?

```{r 1.3}
any(x > 5)

```
1.4) Use a function to see if all the elements of x are greater than 5?

```{r 1.4}
all(x>5)

```
### 2

Given x <- c(10,12,7,2,0,1, NaN, NA), what is the mean(x) after NA is removed? (5 points)

```{r 2}
x <- c(10,12,7,2,0,1, NaN, NA)
mean(x, na.rm = TRUE)

```
### 3

Use a data frame t as the following: (15 points)
set.seed(12345)
t <- data.frame(unit = LETTERS[1:3],  a = rnorm(3),  b = rnorm(3))
class(t)
```{r 3}
set.seed(12345)
t <- data.frame(unit = LETTERS[1:3],  a = rnorm(3),  b = rnorm(3))
class(t)
```
3.1) Create two new columns: sort_a and sort_b by using sort() function of a and b columns.
```{r 3.1}
sort(t$a)

sort(t$b)

```

3.2) Create two new columns: order_a and order_b by using order() function of a and b columns.
```{r 3.2}
order(t$a)

order(t$b)
```

3.3) Create two new columns: rank_a and rank_b by using rank() function of a and b columns.

```{r 3.3}
rank(t$a)

rank(t$b)

```

### 4

Use AutoClaims dataset: (36 points)

4.1) Use read_csv to import AutoClaims data into R
```{r 4.1}
dataPath <- "~/Desktop/MSc Analytics/Autumn 2020/MSCA 37010 2 - Programming for Analytics/Week 3/Assignment/"
dat <- read.csv(paste(dataPath,'AutoClaims.csv',sep = '/'), header=TRUE)
nrow(dat)
head(dat)

```
4.2) Check and remove duplicates

```{r 4.2}
no_dup_dat <- unique(dat)
nrow(no_dup_dat)
head(no_dup_dat)

```
4.3) Sort by Column “Paid”, ascending

```{r 4.3}
library(dplyr)
head(arrange(no_dup_dat,no_dup_dat$PAID))

```
4.4) Sort by Age, descending

```{r 4.4}
head(arrange(no_dup_dat, desc(no_dup_dat$AGE)))

```
4.5) Export the sorted data to a csv file

```{r 4.5}
write.csv(arrange(no_dup_dat, desc(no_dup_dat$AGE)),file=paste(dataPath,'myAutoClaims.csv',sep = '/'))

```
4.6) Subsetting Age>= 90 and Gender = "F"

```{r 4.6}
head(subset(no_dup_dat, AGE >=90 & GENDER == 'F'))

```
4.7) Subsetting Age>= 90 or Gender = "F"

```{r 4.7}
head(subset(no_dup_dat, AGE >= 90 | GENDER == 'F'))

```
4.8) Subsetting PAID >=2000 without Class column

```{r 4.8}
head(subset(select(no_dup_dat,STATE, GENDER, AGE, PAID), PAID >= 2000))

```
4.9) Calculate Average Age by Gender = F and M 

```{r 4.9}
aggregate(x = no_dup_dat$AGE, by = list(no_dup_dat$GENDER),FUN = mean) #Using 'aggregate' function

no_dup_dat %>% group_by(GENDER) %>%  summarise(mean(AGE)) #Using 'dplr' function

```
4.10) Calculate Average PAID by Gender = F and M 

```{r 4.10}
aggregate(x = no_dup_dat$PAID, by = list(no_dup_dat$GENDER),FUN = mean) #Using 'aggregate function'

no_dup_dat %>% group_by(GENDER) %>%  summarise(mean(PAID)) #Using 'dplr' function

```
4.11) Calculate Average PAID by STATE 

```{r 4.11}
aggregate(x = no_dup_dat$PAID, by = list(no_dup_dat$STATE),FUN = mean) #Using 'aggregate' function

no_dup_dat %>% group_by(STATE) %>%  summarise(mean(PAID)) #Using 'dplr' function

```
4.12) Calculate Average PAID by CLASS 

```{r 4.12}
aggregate(x = no_dup_dat$PAID, by = list(no_dup_dat$CLASS),FUN = mean) #Using 'aggregate' function

no_dup_dat %>% group_by(CLASS) %>% summarise(mean(PAID)) #Using 'dplr' function

```

### 5

Merge two datasets by an ID variable, where ID is not the same for both datasets (10 points)
data1 <- data.frame(ID= letters[1:5], x=1:5)
data2 <- data.frame(ID= letters[6:10], y=4:8)

5.1) Keep all the ID from either data1 or data2
```{r 5.1}
data1 <- data.frame(ID= letters[1:5], x=1:5)
data2 <- data.frame(ID= letters[6:10], y=4:8)

merge(data1,data2,all=TRUE)

```

5.2) Keep the ID which are in both datasets
```{r 5.2}
merge(data1,data2)

```

5.3) Only keep the rows from the 1st argument data1
```{r 5.3}
merge(data1,data2,all.x = TRUE)
```
5.4) Only keep the rows from the 2nd argument data2
```{r 5.4}
merge(data1,data2,all.y = TRUE)
```

### 6

Merge two datasets to add columns, where the ID variable has a different name (5 points)
data1 <- data.frame(ID1=1:5, x=letters[1:5])
data2 <- data.frame(ID2=1:5, x=letters[6:10])

```{r 6}
data1 <- data.frame(ID1=1:5, x=letters[1:5])
data1
data2 <- data.frame(ID2=1:5, x=letters[6:10])
data2

merge(data1, data2,  by.x = "ID1", by.y = "ID2")

```

### 7) Merge two datasets to add rows, where the ID variable has a different name (10 points)
data1 <- data.frame(ID1=1:5, x=letters[1:5]);data1
data2 <- data.frame(ID2=1:5, x=letters[6:10]);data2
# Hint: you can rename one of the ID variable so that both ID variables have a same name
```{r 7}
data1 <- data.frame(ID1=1:5, x=letters[1:5]);
data1
data2 <- data.frame(ID2=1:5, x=letters[6:10]);
data2

merge(data1, data2,   all=T)
```
### 8
Use as.Date to convert the following date format to “yyyy-mm-dd” (10 points)
8.1)	dates.1 <- c("5jan2008", "19aug2008")

```{r 8.1}
dates.1 <- c("5jan2008", "19aug2008")
as.Date(dates.1, format="%d%b%Y")

```
8.2)	dates.2 <- c("5-1-2008", "19-8-2008")
```{r 8.2}
dates.2 <- c("5-1-2008", "19-8-2008")
as.Date(dates.2, format="%d-%m-%Y")

```

### 9
(10 points)

9.1) Use as.Date to add 14 days to 2011/1/1 
```{r 9.1}
jan1 <- as.Date("2011/1/1")
(jan8 <- jan1 + 14)
```
9.2) Use as.Date to Subtract 1 week from 2011/2/8
```{r 9.2}
feb <- as.Date("2011/2/8")
(jan8 <- feb - 7)
```
9.3) Compare dates: jan15 > jan1 
```{r 9.3}
jan1 <- as.Date("2011/1/1")
jan15 <- as.Date("2011/1/15")
jan15 > jan1
```
9.4) Use format.Date to extract Year of 2011/1/1
```{r 9.4}
format.Date(jan1, "%Y")
```
9.5) Use format.Date to extract moth and day of 2011/1/1
```{r 9.5}
format.Date(jan1, "%b-%d")
```